Slitasje og kjørespor

### 2022-12-16 DRAFT_slitasjeOgKjørespor.Rmd
infrastruktur index load

#### Infrastructure index

>>>>

#### Outline of Norway and regions

```{r}
outline <- sf::read_sf("data/outlineOfNorway_EPSG25833.shp")
```

```{r}
regions <- sf::read_sf("data/regions.shp")
```

#### Infrastructure index
The infrastructure index is explained [here](https://brage.nina.no/nina-xmlui/handle/11250/2983607).
I use it as a kind of pressure or human influence index. It is likely to be a good predictor for *slitasje*.

```{r}
# The original data is here, on the R server, but it is not always working, so I also include a copy on the P server.

#path <- ifelse(dir == "C:", 
#        "R:/GeoSpatialData/Utility_governmentalServices/Norway_Infrastructure_Index/Original",
#        "/data/R/GeoSpatialData/Utility_governmentalServices/Norway_Infrastructure_Index/Original")
#

path <- ifelse(dir == "C:", 
        "P:/41201785_okologisk_tilstand_2022_2023/data/infrastrukturindeks",
        "/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/infrastrukturindeks")

```{r, eval=F}
infra <- stars::read_stars(paste0(path, "/NY_INFRA_IND.tif"))
```

It's easuer to see what's happeing if we zoom in a bit. 
Lets get a boundary box around Trondheim.
```{r}
myBB <- st_bbox(c(xmin=260520.12, xmax = 278587.56,
                ymin = 7032142.5, ymax = 7045245.27),
                crs = st_crs(infra))
```

Cropping the raster to speed up processing.
```{r}
infra_trd <- sf::st_crop(infra, myBB)
```

Transform to lat long dye to osm requirements
```{r, warning=F}
infra_trd_ll <- sf::st_transform(infra_trd, 4326)
```

Get the boundary box of the cropped raster
```{r}
myBB_ll <- sf::st_bbox(infra_trd_ll)
```

Download highways using the above bbox.
```{r, warning=FALSE, message=F, eval=F}
hw <- 
  osmplotr::extract_osm_objects(
    bbox = myBB_ll,
    key = "highway",
    sf = T)
```

Transforming highway data back into utm, although not strictly necessary.
```{r, eval=F}
hw_utm <- sf::st_transform(hw, sf::st_crs(infra_trd)) 
```

This object contains too many roads. I'll take out the unnamed roads. 
```{r, eval=F}
hw_utm <- hw_utm[!is.na(hw_utm$name),]
```

Caching the sf object
```{r, eval=F}
saveRDS(hw_utm, "data/cache/highways_trondheim.rds")
```

Load object
```{r}
hw_utm <- readRDS("data/cache/highways_trondheim.rds")
```

```{r, fig.cap="Infrastrcture indeks over Trondheim. Major roads are in white."}
tm_shape(infra_trd)+
  tm_raster(title="Infrastructure Index",
            style="cont",
            palette = "viridis")+
  tm_layout(legend.outside = T)+
  tm_shape(hw_utm)+
  tm_lines(col="white")
```

The infrastructure index is in EPSG32633. I will need to transform it. It takes a couple of minutes, so I will store the result in P and load it back in afterwards.
```{r, eval=F}

path <- ifelse(dir == "C:", 
        "P:/41201785_okologisk_tilstand_2022_2023/data/infrastrukturindeks",
        "/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/infrastrukturindeks")

infra <- sf::st_transform(infra, 25833)
stars::write_stars(infra, paste0(path, "/infrastrukturindeksEPSG_EPSG35833.tif"))
```

```{r}
infra2 <- stars::read_stars(paste0(path, "/infrastrukturindeksEPSG_EPSG35833.tif"))
```
